### 2010-07-09 Thesis/Appendix1.tex

\chapter{Appendix Example}
\label{chap:appendix1}

\section{Appendix example section}

And I cite myself to show by bibtex style file (two authors) \citep{Commowick_MICCAI_2007}.

This for other bibtex stye file : only one author \cite{Oakes_RStat_1999} and many authors \cite{Guimond_CVIU_2000}.