### 2010-07-09 Thesis/Resume.tex

\chapter*{Résumé}

\large\textbf{Simulation de tissus mous en temps réel sur GPU pour la simulation médicale} \\

Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla. Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.
\\
Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla. \\

Blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla. \\

Blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla, blablablablabla bla blabla bla blablablabla bla.Blablabla bla blabla blablablabla.