### 2010-07-12 Thesis/Appendix1.tex
Delete most of dummy contents
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@123 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Appendix Example}
\label{chap:appendix1}

\section{Appendix example section}