### 2010-07-14 Thesis/Chapter1.tex
First table of contents + update of bibliography
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@125 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Title}
\label{chap:intro}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}