### 2010-07-14 Thesis/TitlePage.tex
Added draft features, watermark, time stamp
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@126 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{titlepage}
\begin{center}
\noindent {\LARGE \textbf{UNIVERSITY OF LILLE 1}} \\
\vspace*{0.3cm}
\noindent {\LARGE \textbf{DOCTORAL SCHOOL OF ENGINEERING SCIENCE}} \\
\vspace*{0.5cm}
\noindent \Huge \textbf{P H D\ \ T H E S I S} \\
\vspace*{0.3cm}
\noindent \large {to obtain the title of} \\
\vspace*{0.3cm}
\noindent \LARGE \textbf{PhD of Science} \\
\vspace*{0.3cm}
\noindent \Large of the University of Lille 1 \\
\noindent \Large \textbf{Specialty: \textsc{Computer Science}}\\
\vspace*{0.4cm}
\noindent \large {Defended by\\}
\noindent \LARGE Olivier \textsc{Comas} \\
\vspace*{0.8cm}
\noindent {\Huge \textbf{Real-time Soft Tissue Modelling on GPU for Medical Simulation}} \\
\vspace*{0.8cm}
\noindent \Large Thesis Advisor: Stéphane \textsc{Cotin} \\
\vspace*{0.2cm}
\noindent \Large prepared at \\ 
INRIA Lille, \textsc{Shaman} Team and CSIRO ICT Brisbane, EAHRC \\
\vspace*{0.7cm}
%\noindent \large defended on the $00^{th}$ of December 2010 \\
\noindent \large \color{red}{DRAFT - \today ~at \currenttime} \\
\end{center}
\noindent \large \textbf{Jury:} \\
\begin{center}
\noindent \large 
\begin{tabular}{llcl}
      \textit{Reviewers:}	& Bernard \textsc{}		& - & INRIA (Shaman)\\
				& Bernard \textsc{}		& - & INRIA (Shaman)\\
      \textit{Advisor:}	& Stéphane \textsc{Cotin}		& - & INRIA (Shaman)\\
      \textit{President:}	& Bernard \textsc{}		& - & INRIA (Shaman)\\
      \textit{Examinators:}   & Bernard \textsc{}          & - & INRIA (Shaman)\\
      				& Bernard \textsc{}			& - & INRIA (Shaman)\\
      				& Bernard \textsc{}		& - & INRIA (Shaman)\\
\end{tabular}
\end{center}
\end{titlepage}
\sloppy

\titlepage