### 2010-07-30 Thesis/Chapter3.tex
- Update 30 July - Changed outline according to what we discussed with Stephane + increased level of details + bib updated + creation of chapter 3
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@138 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Main principles of Finite Element Method (most used technique to solve equations of continuum mechanics)}
\label{chap3}
\begin{shortAbstract}
Blabla
\end{shortAbstract}

\section{Introduction}

\section{Discretisation}

\section{Derivation of element equations}

\section{Assembly of element equations}

\section{Solution of global problem}