### 2010-08-09 Thesis/Chapter1.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Medical simulation}
\label{chap1}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{General context}
	\subsection{Simulators for everything: we started by simulating what is simple}
	\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}

\section{Challenges}					
	\subsection{Photorealistic rendering}
	\subsection{Accurate haptic feedback}
	\subsection{Physically realist organ modelling in real-time}