### 2010-08-09 Thesis/Chapter10.tex

\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
\label{chap10}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{State of art: reconstruction/simplification}

\section{Our method}