### 2010-08-09 Thesis/Chapter11.tex

\chapter{Summary}
\label{chap11}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}