### 2010-08-09 Thesis/Chapter5.tex

\chapter{Introduction of TLED}
\label{chap5}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}

\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}