### 2010-08-09 Thesis/Chapter8.tex

\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
\label{chap8}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Colon is big, we need a fast model}

\section{Limitations (difficult to link it to actual physical properties)}