### 2010-08-09 Thesis/Chapter9.tex

\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
\label{chap9}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Model description}

\section{Validation}

\section{Application to implant deployment simulation in cataract surgery}