### 2010-08-11 Thesis/Chapter8.tex
- Update 11 August -
Added free-form deformation, shape matching and mass-spring systems + big update for bib



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@149 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Mass-spring model for colon implemented on GPU}
\label{chap8}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Colon is big, we need a fast model}

\section{Limitations (difficult to link it to actual physical properties)}