### 2010-08-17 Thesis/Chapter7.tex

\chapter{State of art: hollow structures}
\label{chap7}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
	\subsection{Colonoscopy simulator project: needs for colon}
	\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}

\section{Non-physic approaches (computer graphics stuff)}

\section{Physically accurate approches (plates/shells)}

% Co-rotational methods
% More recently, \cite{Allard09} modelled the membrane of the lens capsule with a co-rotational finite element formulation in a cataract surgery simulator. The goal was to simulate the anisotropic fracture progation