### 2010-08-25 Thesis/Chapter6.tex

\chapter{GPU implementation of TLED}
\label{chap6}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Re-formulation of the algorithm for its Cg implementation}

\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}

\section{Implementation into SOFA}
	\subsection{Presentation of SOFA project and architecture}
	\subsection{CUDA description}
	\subsection{GPU implementation of TLED in SOFA and release in open-source}

\section{Results}	
	\subsection{Pure shear of a cube}
	\subsection{Compression of a cube}
	\subsection{GPU performance}
	\subsection{Simulation of liver deformation}
	\subsection{A medical application: cataract surgery}

\section{Discussion}
	\subsection{Critical time step}
	\subsection{Difficulties to handle contacts}