### 2010-08-31 Thesis/Chapter7.tex

\chapter{State of art: hollow structures}
\label{chap7}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Introduction: the problematic}
Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?

	\subsection{Colonoscopy simulator project: needs for colon}

	\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}

\section{Non-physic approaches (computer graphics stuff)}

	\subsection{Mass-spring models}

	\subsection{Modal analysis (modal warping)}

\section{Physically accurate approches (plates/shells)}

% Co-rotational methods
% More recently, \cite{Allard09} modelled the membrane of the lens capsule with a co-rotational finite element formulation in a cataract surgery simulator. The goal was to simulate the anisotropic fracture progation