### 2010-09-08 Thesis/Appendix1.tex
- Update 8 September -
Started writing of state of the art for shells + bib updated



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@170 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Tensors \TODO{appendix on tensors necessary?}}
\label{chap:appendix1}