### 2010-09-11 Thesis/Chapter8.tex
- Update 11 September -
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@173 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{\TODO{Mass-spring + mapping for colon on GPU?}}
\label{chap8}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Colon is big, we need a fast model: mass-spring on coarse mesh+ mapping for high res mesh}

\section{Limitations... But difficult since it was somewhat sufficient for the colonoscopy simulator.}