### 2010-10-02 Thesis/Chapter1.tex
- Update 2 October -
A few changes in shell part + part of conclusion




git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@180 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Medical simulation}
\label{chap1}

\section{General context}

	\subsection{Historical evolution: mannequins (plain and electronic), conputer-based simulators}
	\subsection{Ultimate goal: needs to integrate patient-specific + physiology + pathologies for medical training, patient-specific planning and per-operative guidance}

\section{Challenges for computer-based simulators}					
	\subsection{Photorealistic rendering}
	\subsection{Accurate haptic feedback}
	\subsection{Physically realist organ modelling in real-time}