### 2010-10-05 Thesis/Abstract.tex
- Update 5 October -
Added conclusion, abstract, resume and beginning of introduction



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@181 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter*{Abstract}

\begin{center}
\large\textbf{Real-time soft tissue modelling on GPU for medical simulation\\}
\end{center}

\noindent \large\textbf{Abstract:}
Modelling the deformation of anatomical structures in real-time is a crucial problem in medical simulation. Because their shape and their constitution greatly differ, a unique model cannot deal with the variousness of the mechanical behaviours. Hence, we identified two major types of structures in human body, they can be either solid (brain, liver, prostate etc.) or hollow (colon, blood vessels, stomach etc.).

Our answer to this problematic is twofold. Our first contribution is an efficient GPU implementation of a non-linear, anisotropic and viscoelastic finite element model for solid organs. Our second contribution is a framework for real-time modelling of thin anatomical structures via a parallelisable co-rotational shell finite element formulation and a method to mesh a complex surface with curved shell elements.

Although the two soft tissue models are based on continuum mechanics for greater accuracy, they may both be employed to simulate the deformation of entire organs in real-time. Finally, their implementation into the open source framework SOFA will provide worldwide researchers with new models to assist in enhancing the realism of medical simulators.

\bigskip

\large\textbf{Keywords:}
Medical simulation, soft tissue modelling, finite element method, GPU, TLED, shell, co-rotational méthod