### 2010-10-22 Thesis/TitlePage.tex

\thispagestyle{empty}

\begin{center}

\noindent
\begin{tabular*}{\textwidth}{p{5cm}p{8.2cm}}
\multirow{5}{*}{\includegraphics[width=5cm]{images/univ-lille}} & \hspace{\stretch{1}} \\
& \hspace{\stretch{1}} \\
& \hspace{\stretch{1}} \\
& \hspace{\stretch{1}} Thesis submitted to obtain the title of\\
& \hspace{\stretch{1}} Doctor of Philosophy\\ 
\end{tabular*}

\vspace*{0.5cm}
\hspace{\stretch{1}} Doctoral School of Engineering Science\\
\hspace{\stretch{1}} Field: Computer Science \\

\vspace*{2.5cm}
\Huge \textbf{Real-time Soft Tissue Modelling on GPU for Medical Simulation} \\
\vspace*{2.5cm}
\noindent \large Prepared by Olivier \textsc{Comas} at \\ 
INRIA Lille, \textsc{Shaman} team and CSIRO ICT Brisbane, AEHRC \\
\vspace*{0.5cm}
\noindent \large \color{red}{DRAFT - \today ~at \currenttime} \\
%\noindent \large Defended on the $00^{th}$ of December 2010 \\
\end{center}
\vspace*{1cm}
\noindent \large \textbf{Jury:} \\
\begin{center}
\noindent \large 
\begin{tabular}{llcl}
      \textit{Advisor:}	& Stéphane \textsc{Cotin}		& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{Reviewers:}	& Thomas Sangild \textsc{S\o rensen}          & - & University of Aarhus, Denmark\\
				& Georges \textsc{Dumont}		& - & ENS Cachan, Rennes, France\\
      \textit{Examiners:}   & Tamy \textsc{Boubekeur}          & - & Telecom ParisTech, Paris, France\\
      				& Christian \textsc{Duriez}			& - & INRIA (\textsc{Shaman} team), Lille, France\\
      \textit{President:}	& Laurent \textsc{Grisoni} & - & INRIA (\textsc{Mint} team), Lille, France \\
      \textit{Invited:}   & Hans \textsc{De Visser}          & - & CSIRO, Brisbane, Australia\\
\end{tabular}
\end{center}
\sloppy