### 2012-01-04 Doc/Miccai/miccai2012.tex
add Miccai 2012 paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@281 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage[colorlinks=true,linkcolor=black,citecolor=black,pdfnewwindow=true,urlcolor=blue,filecolor=blue,baseurl=./]{hyperref}
\usepackage{url}

% \newcommand{\mycite}[2]{\url{#1}{\cite{#2}}}
% \newcommand{\mycite}[2]{\url{#1}:\cite{#2}}
% \newcommand{\mycite}[2]{\href{http://twitter.com/home}{Twitter}.}
%\newcommand{\mycite}[2]{\href{./refs/#1}{\cite{#2}}}
\newcommand{\mycite}[2]{\cite{#2}}
%\newcommand{\myciteurl}[2]{\href{#1}{\cite{#2}}}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
%\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
%\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
%\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
%\newcommand{\keywords}[1]{\par\addvspace\baselineskip
%\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{color}
\usepackage{amsmath}
\usepackage[tight]{subfigure}
\usepackage{listings}
\definecolor{darkgray}{rgb}{0.2,0.2,0.2}

\lstset{language=c++,basicstyle=\scriptsize\sffamily,tabsize=2,framexleftmargin=5mm, frame=shadowbox, framesep=3pt,rulesepcolor=\color{darkgray},rulesep=.100pt,keywordstyle=\bf\color{blue},commentstyle=\color{magenta},stringstyle=\color{red},numbers=left,numberstyle=\tiny,numbersep=10pt,breaklines=true,showstringspaces=false,columns=flexible}

\newcommand{\tomas}[1]{{\color{red}{\textbf{Tom: #1}}}}
\newcommand{\stefan}[1]{{\color{magenta}{\textbf{St: #1}}}}
\newcommand{\christian}[1]{{\color{green}{\textbf{Ch: #1}}}}
\newcommand{\tobias}[1]{{\color{bleu}{\textbf{Tob: #1}}}}

\newcommand{\myvec}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\mymat}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\Vx}{\myvec{x}} % position vector
\newcommand{\Vv}{\myvec{v}} % velocity vector
\newcommand{\Va}{\myvec{a}} % acceleration vector
\newcommand{\Vdv}{\myvec{dv}} % change of velocity vector (unknown in implicit CG, and used in constraint solver

% \newcommand{\mysubsection}[1]{\noindent \textbf{#1: }}
\newcommand{\mysubsection}[1]{\textbf{#1: }}
% \newcommand{\mysubsection}[1]{\subsection{#1}}

%\usepackage{miseenpage}

\begin{document}

\mainmatter            % start of your contributions

\title{Towards planning of infantile cardiac surgery based on simulation}

\author{PaperID:}

\maketitle

\begin{abstract}
In this paper...
\end{abstract}

\section{Introduction}

\christian{We can put any comment...}
\tomas{People have different colors}

Position vector: $\Vx$

Matrices are written as $\mymat{M}$

\bibliographystyle{abbrv}
\bibliography{miccai2012}

\end{document}