### 2012-02-24 Doc/Miccai/results.tex
moving chapters into separate files to avoid collisions
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@333 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

Example with 3 diff. surgical procedures

Parameters

Limitations (getting image data, meshes, dynamic remeshing, same number of nodes)