### 2012-03-02 Doc/Miccai/conclusion.tex
Improved results and conclusion.
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@385 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

We introduced a new approach for predictive simulation of CHD corrective surgeries. It is based on thin shell elements and a relaxation method for joining blood vessels and patches by topologically connecting affected shell elements. Low-level surgical procedures can be performed universally using this method. We showed that we only need a small number of thin shell elements to achieve predictive simulation results. A sophisticated remeshing algorithm is needed to allow for arbitrary incisions.