### 2013-05-30 Doc/Vrst2013/vrst2013.tex
more fixes in the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@578 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

% TODO
% REMEMBER: After having produced the .bbl file, and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.

\documentclass{acm_proc_article-sp}

\usepackage{caption}
\DeclareCaptionType{copyrightbox}
\usepackage{amsmath}
%\usepackage{makeidx}
%\usepackage[pdftex]{graphicx} ... we have to use ps/eps
\usepackage{subfig}

\def\wrt{w.\,r.\,t.}
\def\eg{e.\,g.}
\def\ie{i.\,e.}
\def\Dash{\nobreak\,---\penalty-500\,}

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
\newcommand{\IP}[1]{{\color{cyan}\textbf{IP: #1}}}
\newcommand{\TODO}[1]{{\color{red}\textbf{TODO: #1}}}
%%%

\renewcommand{\Vec}[1]{\mathbf{#1}}
\newcommand{\Mat}[1]{\mathbf{#1}}

\begin{document}

\title{Modeling Liver with Glisson's Capsule in Real-Time Medical Simulations}
%\subtitle{}

\numberofauthors{4}
\author{
% 1st. author
\alignauthor
Tom\'a\v{s} Golembiovsk\'y\\
       \affaddr{Faculty of Informatics}\\
       \affaddr{Masaryk University}\\
       \affaddr{Brno, Czech Republic}\\
       \email{nyoxi@ics.muni.cz}
% 2nd. author
\alignauthor
Igor Peterl\'ik\\
       \affaddr{Institut Hospitalo-Universitaire}\\
       \affaddr{Strasbourg, France}\\
       \email{...} % TODO
% 3rd. author
\alignauthor
Christian Duriez\\
       \affaddr{INRIA}\\
       \affaddr{Lille, France}\\
       \email{...} % TODO
\and
% 4th. author
\alignauthor
Stephan\'e Cotin\\
       \affaddr{INRIA}\\
       \affaddr{Lille, France}\\
       \email{...} % TODO
}

%\date{30 May 2013}

\maketitle
\begin{abstract}
  \TODO{}
The accurate mechanical modeling of liver is of a paramount interest in simulation-based operation planning 
and computer-aided per-operative guidance. However, the realistic simulation of liver behaviour is a challenging task, 
since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, each having different mechanical properties. While the 
mechanical influence of vessels on both global and local behaviour of the organ has been measured and modelled previously~\cite{Peterlik2012}, only local 
effects have been studied in the case of Glisson's capsule~\cite{Hollenstein2006,Ahn2010}.

In this paper we introduce what we believe is the first real-time model of liver taking into account all three constituents, each being simulated 
employing a different finite element formulation.
Using the proposed model, we focus on the role of the Glisson's capsule: first, we show that the model is able to reproduce correctly
the local effects of the capsule in so called aspiration test. Second, based on CT image data of porcine abdominal cavity acquired in supine and flank 
positions, we demonstrate 
that the capsule has also a global impact on the mechanical behaviour of the liver in the case when the organ undergoes large deformations.

%This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
%Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
%Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
%However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
%In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver.
%We show that the approach is able to reproduce the deformations observed at the local level.
%Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

\end{abstract}

% TODO
% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory} %TODO

% NOT required for Proceedings
%\keywords{deformable liver model, Glisson's capsule, real-time soft tissue modeling}
\keywords{deformable liver model}

\section{Introduction} %{{{

According to the statistics, nearly 100,000 European citizens die of cirrhosis or liver cancer each year. 
Although new methods such as radio-frequency- and cryo-ablation known in the interventional radiology 
seem to be promising, surgery remains the option that offers the foremost success rate against these pathologies. 
Nevertheless, surgery is not always performed due to several limitations, in particular the determination 
of accurate eligibility criteria for the patient. 
In this context, the pre-operational planning becomes a crucial task having a significant impact on the treatment.

Computer-aided physics-based medical simulation has proven to be an extremely useful technique in the area of medical training. 
Nevertheless, whereas generic models are usually required in training simulators, accurate patient-specific modelling
becomes necessary as soon as computer simulation is to be employed in the pre-operative planning. At the same time, 
interactivity of such models remains an important aspect, requiring real-time simulation which is often difficult to 
achieve given the complexity of soft tissues.

When simulating the behavior of human liver, the task of real-time accurate modeling is challenging, mainly because of the complex structure 
of this organ composed of three main constituents: \emph{parenchyma}, \emph{vascular networks} and \emph{Glisson's capsule}.
The parenchyma has certainly been the most studied component of the liver; actually researchers agree on hyperelastic 
properties of the tissue, for which the mechanical parameters have been reported for example in~\cite{Kerdok2006,Gao2009}. 
Moreover, several methods have been proposed to model the hyperelastic behaviour at real-time rates, such as multiplicative Jacobian decomposition
introduced in~\cite{Marchesseau2010}.

\TG{The following paragraph is optional. Unnecessary but adds to the global context}
The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that the 
influence of vessels on the mechanical behaviour of the organ is significant. %, mainly if large deformations occur. 
%Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
%the authors propose a composite model allowing for real-time simulation of entire liver with vascularization.
Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
the authors propose a composite model allowing for real-time simulation.

Rather a small number of studies have been conducted dealing with the third liver constituent, the  Glisson's capsule.
Quantitative results of experiments on a porcine liver have been published in~\cite{Umale2011}; the measurements indicate that although being very 
thin (10--20$\mu m$), the capsule shows to be stiff in tensile tests: the Young's modulus of the capsule reported to be $8.22\pm3.42$\,MPa 
exceeds the values for the parenchyma by three orders of magnitude.
This suggests that the mechanical influence of the membrane on the liver behaviour is not negligible.

In~\cite{Hollenstein2006}, a local influence of the capsule has been measured using a special aspiration device. The study was then repeated 
in vivo on human patients during the operation, confirming the mechanical importance of the membrane~\cite{Ahn2010,Nava2008}.
%To our best knowledge, no attempt has been made to demonstrate the role of the Glisson's capsule on the global behaviour of 
%the liver, mainly if the organ undergoes large deformations which is often the case during the surgery.

In this paper we present preliminary results of our work on the complete liver
model. We present a real-time composite model accounting for parenchyma and
Glisson's capsule. The model is based on two different finite element
representations for each constituent coupled together.
%The main contribution of the paper is twofold: first, we present a real-time model of the liver, including
%the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
%linked together via mechanical coupling.
We show that the model mimics the \emph{local} experiments described in~\cite{Hollenstein2006}.

%Second, we use the complete model of the liver to demonstrate the \emph{global} influence of the Glisson's
%capsule via simulation: using a specific model of a porcine liver built from CT contrast-enhanced data, we show that there is a significant 
%difference in the response of the model with and without the capsule in the case when the liver undergoes large deformations.

The paper is organized as follows: first we describe the proposed model of
liver with capsule. Second, we validate our model in context of local
deformations by reproducing the aspiration test described
in~\cite{Hollenstein2006}. We conclude by summarizing future steps towards the
complete liver model.
%Finally, we demonstrate that in spite of its small thickness, the Glisson's 
%capsule has a global influence on the liver undergoing large deformations.

%}}}

\section{Model} %{{{

In this section we describe the construction of a composite model
based on two components: tetrahedral FE model of the 
parenchyma and
triangular membrane elements used for the capsule.

\subsection{Parenchyma} %{{{

It is known that the parenchyma exhibits non-linear viscoelastic behaviour \cite{Marchesseau2010}.
However, as we are mainly interested in the static equilibrium, we do not model the time-dependent
phenomena related to viscosity.

%However, we employ simpler corotational model as we are not so much
%interested in time-dependent behaviour but rather in static equilibrium
%under certain conditions.
%We also rely on the vascularized model of the parenchyma proposed by
%Peterl\'{i}k et al. \cite{Peterlik2012}.

The parenchyma is modeled using corotational finite elements~\cite{Felippa2005}.
Although it relies on linear stress-strain relationship, large displacements including rotations are modelled correctly. 
While in the full non-linear formulation the stiffness matrix relates the forces $\Vec{f}$ and 
displacements $\Vec{u}$ as $\Vec{f} = K(\Vec{u})$, the corotational model 
requires the stiffness matrix $\Mat{K}_0$ of the system to be computed only once before the simulation begins. 
Then, in each step, the motion of each element $e$ is decomposed into rigid rotation $\Mat{R}^e$ and local deformation. 
The rotations are then used to update each local element stiffness matrix as $\Mat{R}\Mat{K}_0\Mat{R}^{\top}$
whereas the deformations are used to compute the linear strain in the local corotational frame.
There are several ways of computing the corotational frame for elements; we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.

%The model of the vascularization is based on linear beams 
%with local frames of reference~\cite{Duriez2006}; in many aspects it's similar to the
%corotational formulation described above. As such the model also handles geometric
%non-linearities in the deformation. Through a specification of cross section and moments of inertia, 
%the model can account for the specific properties of the blood vessels.

%The beam-based model of vessels is mechanically coupled to the parenchyma as described in~\cite{Peterlik2012}. 
%The coupling assumes that there is no relative motion between the vessels and the surrounding parenchyma. 
%During the simulation, the nodes of linked beams are first displaced and rotated according to the actual motion of associated tetrahedra. 
%As the deformation of beams results in mechanical response represented by forces and torques, these are propagated back to 
%the tetrahedral FE model.

%}}}

\subsection{Glisson's Capsule} %{{{
\label{ss:capsuleModel}
The thickness of the Glisson's capsule is relatively small: the values in range of 10-20
$\mu$m have been reported in~\cite{Umale2011}.
It is not possible to model such thin structure with classical tetrahedral
elements, if the real-time aspect of the simulation is to be preserved.
Furthermore, modeling both the tissue and the capsule would require an extremely 
dense mesh to avoid numerical instabilities and thus would significantly
violate the speed requirements imposed for medical simulators.
Instead, modeling the capsule with two-dimensional elements that abstract from the
thickness in the third dimension seems
to be a natural choice. In the elasticity theory, this functionality is usually provided by membrane and shell elements.
Based on the observation of its behaviour, we also
assume negligible bending forces and propose a model based on membrane
elements. 
To maintain simplicity of the composite model we choose simple triangular
elements with constant strain (CST).

The computation of elastic stiffness matrix follows the common derivation
%
\begin{eqnarray}
  \Mat{K}^m & = & \int_V \Mat{B}^T \Mat{E} \Mat{B} dV     \label{mem1} \\
            & = & h \int_A \Mat{B}^T \Mat{E} \Mat{B} dA   \label{mem2} \\
            & = & h A \Mat{B}^T \Mat{E} \Mat{B}           \label{mem3}
\end{eqnarray}
%
where $\Mat{B}$ is the strain-displacement matrix, $\Mat{E}$ the material
matrix, $h$ is the thickness and $A$ area of the element. In the previous
\eqref{mem2} follows from the fact that we assume constant thickness of the
element and \eqref{mem3} follows from the fact that the strain-displacement
matrix is constant in our case. The strain-displacement matrix for the CST
element can be expressed as:
%
\begin{equation}
  \Mat{B} = \frac{1}{2A} \begin{bmatrix}
    y_{23} & 0      & y_{31} & 0      & y_{12} & 0 \\
         0 & x_{32} & 0      & x_{13} & 0      & x_{21} \\
    x_{32} & y_{23} & x_{13} & y_{31} & x_{21} & y_{12}
  \end{bmatrix}
\end{equation}
%
The values $x_{ij} = x_i - x_j$ and $y_{ij} = y_i - y_j$ are computed from
the $x$ or $y$ coordinates of the nodes $i,j$ of the triangular element.
The reader can refer to the respective literature for more thorough
description~\cite{Felippa2003}.

Similarly as with model of parenchyma we use linear elastic material and employ
the corotational formulation for the CST elements.

%}}}

\subsection{Coupling Between Capsule and\\Parenchyma} %{{{
The literature reports high cohesion between capsule and parenchyma.
Based on this property we assume there is no relative motion of the capsule \wrt\ the parenchyma.
Although an arbitrary surface mesh could be used to model the capsule, we exploit 
the fact that the parenchyma is modeled by tetrahedral elements having
triangular faces. Thus, as the boundary of the volumetric mesh is already
triangulated, we simply employ the triangles on the mesh surface to model the capsule.

Using directly the boundary of the tetrahedral mesh does not only solve the
problem of building the surface mesh, but has one more advantage: the nodes
of the triangular mesh coincide with the nodes of the tetrahedral mesh, so no projection of one mesh onto the other is needed.
Moreover, the stiffness matrices for capsule and parenchyma are then easily assembled
together.
% and solved as one system.
%
%\CD{The following may appear as "trivial"... maybe we can remove this part at the end if we need space} 
%Without the loss of generality we can assume the tetrahedron consists of
%nodes $p_1, p_2, p_3$ and $p_4$ and the boundary triangle has nodes $p_1, p_2$
%and $p_3$. We can reorder the degrees of freedom so that the stiffness
%matrix $\Mat{K}^t$ for the tetrahedron can be written as:
%%
%\begin{equation}
%  \Mat{K}^t = \left[\begin{array}{c|c}
%      \Mat{K}^t_{1-3,1-3} & \Mat{K}^t_{1-3,4} \\
%      \hline
%      \Mat{K}^t_{4,1-3} & \Mat{K}^t_{4,4} \\
%  \end{array}\right]
%\end{equation}
%%
%Then the assembled stiffness matrix for the element is:
%%
%\begin{equation}
%  \Mat{K} = \left[\begin{array}{c|c}
%      \Mat{K}^t_{1-3,1-3} & \Mat{K}^t_{1-3,4} \\
%      \hline
%      \Mat{K}^t_{4,1-3} & \Mat{K}^t_{4,4} \\
%  \end{array}\right]
%  +
%  \left[\begin{array}{c|c}
%      \Mat{K}^m & 0 \\
%      \hline
%      0 & 0 \\
%  \end{array}\right]
%\end{equation}
%%
%Where $\Mat{K}^m$ is the stiffness matrix of the triangular membrane.
The resulting system of linear equations is solved by direct solver based on Cholesky decomposition.

Alternatively, one can use a mechanical coupling similar to the one used in
\cite{Peterlik2012} to be able to use an arbitrary surface mesh. Nevertheless,
for conforming triangular mesh both methods lead to the same solution.

%}}}

%}}}

\section{Aspiration Test} %{{{

The model presented in the previous section was implemented in 
SOFA\footnote{www.sofa-framework.org} and a set of
numerical simulations was performed.
In this section we provide comparison of local deformations with the
results reported in literature to validate the method.
%Second, to show that in spite of its very small thickness the membrane cannot be
%neglected even in the context of global deformations and its overall
%stiffness plays an important role, the model of
%the complete liver was subjected to global deformations.

During the contact with an instrument such as probe, needle, scalpel and others,
a specific deformations take place in the vicinity of
the instrument. This type of deformation may not necessarily induce the
deformation of the object as a whole and therefore can be considered as
local. Correct material properties are not only important to quantify the
displacement, but also play an important role in capturing the correct area of
the deformation or its profile near the instrument.

\begin{figure}
  \centering
  \includegraphics[height=3.5cm]{aspiration.jpg}
  \caption{\label{fig-aspiration1} The SOFA simulation scene of the aspiration test.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=7cm]{aspiration.pdf}
  \caption{\label{fig-aspiration2} Displacement profile of the cube in the
  aspiration test with (blue) and without (black) the capusle.}
\end{figure}

A good example of such a local deformation is the aspiration test
where the response of liver exposed locally to a negative pressure is measured.
The aspiration device consists of a tube 1\,cm in diameter and allows
control of internal pressure in the tube. The test is performed by
attaching the tube to the tissue and measuring the tissue response. We
set up a simulation in SOFA to reproduce the experiment (see
Fig.~\ref{fig-aspiration1}: we meshed a 15$\times$15$\times$15\,mm$^3$ 
cube representing the tissue 
resulting in 2648 tetrahedra. Then we attached a 1\,cm tube 
and applied a pressure of 3\,kPa inside the tube. The interaction between the tube and 
tissue was modelled with friction contact.

\TG{Here we have to decide wether to keep the values below as well as the Fig.
2 and specify the values used to match Hollenstein separately. Or if we take
just one set of values and redo the figure.}

In the simulations we used material properties for a pig liver reported in the literature: for the 
Glisson's capsule we used Young's modulus $E_c$=8.22\,MPa and thickness $t_c$=20\,$\mu$m~\cite{Umale2011}.
Since the values for parenchyma elasticity quite vary (usually between 2\,kPa to 5\,kPa), we perform the simulation 
with $E_p$=3.5\,kPa.

In~\cite{Hollenstein2006} it is reported that the Glisson's
capsule has a non-negligible influence on the response of the tissue and modeling
only the parenchyma leads to an overestimation of the deformation up to the factor of 3. 
In Fig.~\ref{fig-aspiration2} the profiles of cuts in the middle of the
test cube are presented showing significantly larger deformation in the model without capsule.
We can observere that the deformation without the capsule is indeed overestimated approximately by the factor of 2. This is in good agreement with the results reported in~\cite{Hollenstein2006}

Moreover, if we neglect the time-dependent behaviour we are able to reproduce
the measurements presented in~\cite{Hollenstein2006,Nava2008} with less than 1\% error.
In this case the simulation was performed with the pressure of 30\,kPa and with
the following elastic parameters: $E_p$=30\,kPa for parenchyma and $E_c$=3\,MPa
and $t$=93\,$\mu$m for capsule. All the values lie within the measurements
specified in the paper. \TG{add figure like Fig. 2?}

%and it can be concluded that 
%our composite model based on coupling the triangular and tetrahedral elements provides an accurate 
%model of parenchyma and capsule.

%}}}

%\begin{figure*}
%  \centering
%    \includegraphics[width=.95\textwidth]{forceEvolution.pdf}
%  \caption{\label{f:forceEvol}Evolution of forces during the simulation, where the supine data is deformed
%towards the flank shape, and the stiffness of the capsule is increased from 0 to 8.22\,MPa after the initial equilibrium is achieved.}
%\end{figure*}

%\subsection{Global Deformations}
%In the previous section we have shown that the composite model of parenchyma and 
%capsule can be used to reproduce the \emph{local} behaviour of the tissue. In the following, 
%we first present a real-time complete model of entire liver built from image data. 
%Second, using this model we demonstrate the \emph{global} mechanical influence of the capsule when 
%the model undergoes large deformations. 
%
%The complete model was constructed using contrast-enhanced CT data acquired on a female pig 
%being positioned first in supine and then flank position. 
%The liver and hepatic vein were segmented for both datasets  using semi-automatic methods available in ITKSnap~\footnote{www.itksnap.org}.
%Further, an unstructured mesh was obtained from the segmented maps using CGAL~\footnote{www.cgal.org}. 
%The semi-automatic method presented in~\cite{Peterlik2012} was 
%then applied to the segmented map of the hepatic vein, resulting in linked-beam model. 
%Finally, the model of the Glisson's capsule was constructed using the elements described in 
%section~\ref{ss:capsuleModel} over the surface triangles extracted from the tetrahedral mesh. 
%The complete model composed of 4266 tetrahedra, 188 beams and 1284 triangles was then 
%used in simulation where external force representing the gravity was applied to the model, 
%resulting in large deformations mainly in the area of lobes. The refresh rate of 22\,FPS 
%was achieved on a desktop computer with Intel i7 CPU running at 3.4\,GHz and 8\,GB of memory. 
%The simulation with gravity was also used as a first demonstration of the global influence of the capsule. 
%In Fig.~\ref{f:global}a, the models with capsule (red) and without (yellow) are depicted after being 
%deformed due to the gravity, showing a significant difference between the two models.  
%
%\begin{figure*}
%  \centering
%    \subfloat[]{\includegraphics[width=.32\textwidth]{gravity2.png}}
%    \subfloat[]{\includegraphics[width=.32\textwidth]{registered.png}}
%    \subfloat[]{\includegraphics[width=.32\textwidth]{distanceMap.png}}
%    \caption{\label{f:global}%
%      (a) Liver under gravity with (red) and without (yellow) capsule;
%      (b) liver in supine (yellow) and flank (red) position;
%      (c) Hausdorff distance map between registered version with and without capsule.}
%\end{figure*}
%
%In order to demonstrate the influence of the capsule on a real deformation of the liver, we 
%performed a following protocol employing the supine and flank models built from the CT data. 
%First, we performed a rigid registration of the two organs \wrt\ the root point of the hepatic vein. 
%As shown in Fig.~\ref{f:global}b, the liver undergoes very large deformation when the body 
%is turned from the supine to the flank position, exceeding 6\,cm for some parts of the organ. 
%In order to induce similar deformation, we extracted feature pairs from the vascular system of the liver:
%we selected points such as bifurcations which can be identified easily in both supine and flank 
%data resulting in a set of correspondences. In the next step, we used these correspondences as a non-homogeneous
%boundary conditions (imposed via penalty term) in order to perform a deformable registration of the supine
%model onto the flank data. During the registration, the response forces acting in the points with prescribed displacements
%were calculated. It is important to emphasize that the goal of the simulation was not 
%to perform an accurate registration of the flank and supine data, but the simulation was used 
%as an example of a real deformation of the organ.
%
%The simulation was performed with the model composed of parenchyma,  vessels and capsule, however, zero stiffness was 
%set for the last component.
%Once the static equilibrium was achieved, the effect of the capsule was gradually switched 
%on, \ie\ the stiffness of the membrane was slowly increased from 0 to 8.22\,MPa. The evolution of the force response in the 
%points with prescribed displacements is depicted in Fig.~\ref{f:forceEvol} showing the change in the response as the effect
%of the capsule is included. At the same time, the shape of the deformed liver with capsule was compared 
%to the model without the capsule. A Hausdorff distance map for those cases is depicted in Fig.~\ref{f:global}c: in the red areas, the 
%maximal difference exceeding 1.25\,cm was measured. We believe that this result confirms that the Glisson's capsule 
%plays an important role in the global mechanical behaviour of the liver.

%}}}

\section{Present and Future Work} %{{{

Presently we are working on the validation of the model also in context of
global deformations where we believe the capsule also plays an essential role.
This task is however complicated by the lack of biomechanical literature
dealing with the response of the capsule.
%The influence of the capsule on local level has been published in literature
%before. However, to our knowledge it's influence on global scale deformation
%has not been studied yet. We believe that the capsule has non-negligible impact
%also on global level.

Our goal is to create complete liver model suitable for real-time medical
simulations. For this a model accounting for all three main constituents
(parenchyma, vascularization and capsule) is necessary. To achieve that we have
integrated our model with the vasularization method proposed
in~\cite{Peterlik2012}. In our future work we need to validate the model in the
context of global and local deformations. This task is challenging because
comparing in vivo measurements with the simulation is difficult due to complex
boundary conditions. 
\TG{ex-vivo: no information abou vessels; phantoms: cannot model capsule}

%}}}

\section{Conclusions} %{{{

In the paper we presented a liver model accounting for the Glisson's capsule, a component that is often overlooked. Each constituent being modelled with a different type 
of finite elements: tetrahedral corotational model used for the parenchyma and
and CST membrane elements used for the capsule. 
The model was validated in the context for local displacements based on the the aspiration test presented in~\cite{Hollenstein2006}. The aspiration test
was successfully reproduced, showing a good agreement between the simulation and experiments.

%Further, the model was used to demonstrate global influence of the capsule when the organ 
%undergoes large deformations.

%In the future work, we plan to employ the proposed model for other organs where the capsule play an important role
%(for example kidneys). We will also focus on better description of the mechanical interactions and couplings among the organs and tissues in the abdominal 
%cavity via realistic modeling of boundary conditions using mechanical constraints and contacts.

%}}}

%
% ---- Bibliography ----
%

\bibliographystyle{abbrv}
\bibliography{bibdata}

\end{document}
% spell: setl spell spelllang=en spellfile=spelldict.en.utf-8.add
% vim:set et sw=2 fdm=marker fdl=3 fdc=4 isk+=_,-: