### 2013-07-31 Doc/MMVR2014/figures.tex

\newpage

\begin{figure}
  \centering
  \includegraphics[height=3.5cm]{aspiration.jpg}
  \caption{\label{fig-aspiration1} The SOFA simulation scene of the aspiration test.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=7cm]{aspiration.pdf}
  \caption{\label{fig-aspiration2} Displacement profile of the cube in the
  aspiration test with (blue) and without (black) the capsule.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=3.5cm]{displacement.pdf}
  \caption{\label{fig-aspiration3} Evolution of the displacement at the center
  in time for test with (blue) and without (black) the capsule.}
\end{figure}