### 2013-08-02 Doc/MMVR2014/abstract.tex
FIX: finalization of the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@608 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{frontmatter}%
\begin{abstract}%
Accurate biomechanical modeling of liver is of a paramount interest for pre-operative planning or computer-aided per-operative guidance. 
However, such simulations remain a challenging task, since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, 
each having different mechanical properties. 
%
%While a real-time simulation of vascularized liver has been already described in~\cite{Peterlik2012}, the Glisson's capsule has not been included in the modeling. 
%In~\cite{Hollenstein2006,Ahn2010} it is shown that the capsule plays an important role mainly in the vicinity of the surgical tools. 
%Therefore, the accuracy of the liver response in the vicinity of the surgical tool requires correct modeling of the capsule as a component of the liver model. 
%
In this paper we propose a complete liver model, where the parenchyma is modeled as a finite element volume, the vessels are modeled as series of beam elements, 
and the corotational membrane elements based on constant-strain formulation are used for capsule. These components are coupled mechanically as shown in the paper.
We show that using this model we are able to reproduce accurately biomechanical tests, while maintaining the real-time aspect of the simulation.

%The measures performed on the capsule report stiffness which is significantly higher than that of the parenchyma, however, the thickness of the membrane does not exceed 100$\mu$m. 
%For this reason, it is not possible to model the capsule with standard volume elements usually employed for the parenchyma. 
%In our approach, we rely of corotational membrane elements based on constant-strain formulation, which are coupled mechanically to the underlying tetrahedra.

\end{abstract}%
% 
\begin{keyword}%
Simulation , modeling, liver, biomechanics
\end{keyword}%
% 
\end{frontmatter}%