### 2013-10-28 Doc/MMVR2014/IOSProceedings/conclusion.tex
ADD: proceedings version of the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@610 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\section{Conclusion}
In this paper we presented a composite liver model based on volume and membrane elements. 
When compared to experimental data provided in the literature we found similar results, but with faster computation times than standard computationally expensive model based entirely on volume elements.

%Presently we are focusing on the validation of the model also in context of global deformations of the organ where we believe the capsule plays an essential role as well. 
%This task is however complicated due to the limited number of experimental studies dealing with both the parenchyma and the capsule.

%\section*{Acknowledgments} 
%The authors would like to thank the French Research Agency (ANR) for funding this study through the ACouStiC project.